---
title: "Catologue of Life Taxon cleanup"
author: "Matthew watts"
date: "7 February 2018"
output:
  word_document: default
  html_document: default
---

**This following R script is used to cleanup and organise The Catalogue of Life's database for use with the VectorBite databases. For more information about the The Catalogue of Life database please visit http://www.catalogueoflife.org/. **

In order to run this code in R, you to install and load R's 'Tidyverse' package (see https://www.tidyverse.org/) . 


###Note this currently produces duplicates, need to fix######

```{r}

library(tidyverse)

```



#Load in the cat of life data set
```{r}
taxa <- read_delim("C:/Users/matt/OneDrive - Imperial College London/DataSets/Taxonomy/CatelogueOfLife/archive-kingdom-animalia-bl3/taxa.txt", 
     "\t", escape_double = FALSE, col_types = cols(acceptedNameUsageID = col_character(), 
         datasetID = col_character(), parentNameUsageID = col_character(), 
         taxonID = col_character()), trim_ws = TRUE)

```

```{r}

head(taxa)

```


#filter the cat of life data set to remove all other entries other than accepted name


```{r taxa-select&filter}

taxa_processed <- 
  taxa %>%
    filter(taxonomicStatus == "accepted name")  %>%
  select(taxonID, scientificName, kingdom, phylum, class, order, superfamily, family, genericName, genus, subgenus, specificEpithet, infraspecificEpithet)

taxa_processed <-
  taxa_processed %>%
  cbind(genus2=rep(taxa_processed$genus)) %>%
  cbind(specificEpithet2=rep(taxa_processed$specificEpithet)) %>%
  unite_("species", c("genus2","specificEpithet2"), sep = " ")
  
taxa_processed <- distinct(taxa_processed)
write.csv(taxa_processed, file = "CATLIFE.csv", row.names = FALSE)
#rm(taxa_clean, taxa)

```


```{r}
head(taxa_processed)

```

